
\begin{algorithm}
\caption{Basic Autonomous competence identification}
\begin{algorithmic}[1] % Enables line numbering
\State \textbf{Prerequisites:} Minimal group size $N_{min}$, Time constant $t_{c}$, Participation cost $X_g$, Competence Asset $C_i$, Protocol Treasury $T$, Minimal turn count ${M, M > 2}$
\State \textbf{Process:}
\Procedure{Preparation}{}
    \State Instance is created at timestamp $t_0=t$, with creators personal join requirements $X_a$
    \State Participants join the group and lock in a commitment: $N_{min} \times (X_g+X_a) \rightarrow T$
    \State Joining period ends at $t_1  \ge t_c + t_0$
\EndProcedure
\Procedure{FirstTurn}{}
    \While{$t_1 + t_c \ge t$}
        \If{new Proposal}
        \State Record ${P_i}$
        \State Record $i$
        \EndIf
    \EndWhile
    \If{Group is full or joining timeout}
        \State Members exchange encrypted proposals (MPC).
    \EndIf
\EndProcedure
\Procedure{DecryptProposals}{}
    \If{Timeout or all proposals received}
        \State Members decrypt proposals using threshold encryption.
    \EndIf
\EndProcedure
\Procedure{RegularTurn}{}
    \State Submit new encrypted proposals.
    \State Submit encrypted votes and zk-proof of valid vote.
\EndProcedure
\Procedure{CalculateRoundScores}{}
    \If{Round ends or all votes \& proposals submitted}
        \State Use TSS cryptography to decrypt everything.
        \State Calculate round scores.
    \EndIf
\EndProcedure
\Procedure{RepeatUntilXProposals}{}
    \State Repeat steps 3-5 until a total of X proposals.
\EndProcedure
\Procedure{FinalizeRankings}{}
    \State Count total votes and sort proposals.
\EndProcedure
\end{algorithmic}
\end{algorithm}

